---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Energy Distribution in North Carolina"
subtitle: "https://github.com/meganlundequam/LundequamVanasse_EDA_FinalProject"
author: "Megan Lundequam, Casey Slaught, Sam Vanasse"
fontsize: 12pt
mainfont: Times New Roman

---

RMarkdown Cheat Sheet: https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf


\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
library(scales)
library(sf)
library(tidyverse)
library(plyr)
library(tidycensus)
library(tidyr)
options(tigris_use_cache = TRUE)
library(dplyr)
library(lubridate)
library(corrplot)


# Load your packages

# Set your ggplot theme
theme_set(theme_classic())

# Load your datasets
censusData = st_read("../Data/Processed/NC_county_complete.geojson")
```


# Rationale and Research Questions



\newpage

# Dataset Information



\newpage

# Exploratory Analysis

There is a marked difference in median household income across counties in North Carolina.

```{r, echo=FALSE, message=FALSE, fig.height=9}

ggplot(censusData[censusData$Year==2020,], 
       aes(x=MedianFamilyIncome, y=reorder(CountyShort, MedianFamilyIncome))) +
  geom_point() + 
  scale_x_continuous(labels=comma) + 
  labs(x="Median Household Income", y="NC Counties") + 
  theme(text = element_text(size = 8))

```

Energy generators are diverse in type and widespread throughout the state. Yet while solar appears to be the most abundant, it makes up small fraction of total energy generation.

```{r, echo=FALSE, results='hide', fig.height=7, fig.width=10}

generators = st_read("../Data/Processed/generatorsFuelType.geojson")

ggplot() + 
  geom_sf(aes(fill=MedianFamilyIncome), data=censusData) + 
  geom_sf(aes(color=FuelType), alpha=0.6, size=3, data=generators) + 
  scale_y_continuous(labels=comma) + 
  scale_fill_gradient(labels=comma) + 
  scale_color_manual(values=c(
    "Coal" = "#9c0310",
    "Petroleum Product" = "#e00b8b",
    "Natural Gas" = "#d49a08",
    "Gaseous Renewable Fuel" = "#edc600",
    "Energy Storage" = "#8f8f8f",
    "Liquid Renewable Fuel" = "#aed100",
    "Nuclear" = "#870ee3", 
    "Gaseous Renewable Fuel" = "#778a3f",
    "Hydroelectric" = "#147aff",
    "Solid Renewable Fuel" = "#465e00",
    "Waste Heat" = "#46998d",
    "Solar" = "#00ff7b",
    "Wind" = "#80a6ff"
    )
  ) + 
  labs(color="Generator Type", fill="Median Household Income") + 
  theme_void() 

```


\newpage

# Analysis


```{r include=FALSE}
## Adding new census info and cleaning it
additional.census <- read.csv("../Data/Processed/UpdatedCensus.csv", stringsAsFactors = TRUE) 
additional.census <- 
  additional.census %>%
  subset(select=c(County, TotalPopulation, NoSchooling, Mobility, RaceAllocation, DetailedRace)) %>%
  rename(c("PovertyMobility" = Mobility, "WhitePopMobility" = RaceAllocation, "CollegeDeg" = DetailedRace)) %>%
  mutate(Per.PovertyMobility = (PovertyMobility/TotalPopulation),
         MinorityMobility = (TotalPopulation - WhitePopMobility),
         Per.MinorityMobility = (MinorityMobility/TotalPopulation),
         Per.NoSchooling = (NoSchooling/TotalPopulation),
         Per.CollegeDeg = (CollegeDeg/TotalPopulation)
         )

## Generator Info
census.data <- read.csv("../Data/Processed/generatorsCensusCategoriesComplete.csv", stringsAsFactors = TRUE)
census.data <- 
  full_join(census.data, additional.census) %>%
  drop_na(Energy.Source.1)

## Getting nameplate capacity
generator.data <- read.csv("../Data/2020EnergyData/3_1_Generator_Y2020.csv", stringsAsFactors = TRUE)

## Selecting variables of interest
generator.data <- 
  generator.data %>%
  subset(generator.data$State == "NC") %>%
  subset(select = c(County, Plant.Code, Prime.Mover, Nameplate.Capacity..MW.))

## Adding income to the full generator data by county
full.data <- cbind(census.data, generator.data$Nameplate.Capacity..MW.) %>% 
  rename(c("Nameplate.Capacity" = `generator.data$Nameplate.Capacity..MW.`)) %>% 
  mutate(Nameplate.Capacity1 = as.numeric(as.character(Nameplate.Capacity)))

```


```{r include=FALSE}
## Aggregating the generator data so that it groups all of it by county and by generation type
aggregated.gen.data <- 
  full.data %>%
  group_by(County, Renewable.v.FossilFuel) %>%
  dplyr::summarise(energy.type.gen = sum(Nameplate.Capacity1))

## Joining it to the census data to have capacity and income in same file & selecting specific variables
processed.gen.census <- 
  full_join(aggregated.gen.data, census.data) %>%
  subset(select=c(County, Renewable.v.FossilFuel, energy.type.gen, MedianHouseholdIncome, TotalPopulation, Per.CollegeDeg, Per.PovertyMobility, Per.MinorityMobility, Per.NoSchooling)) %>%
  distinct(.keep_all = TRUE)

## Creating a total generation variable, Removing fossil vs renewable distinction
total.gen.census <- 
  processed.gen.census %>%
  group_by(County) %>%
  dplyr::summarise(total.gen = sum(energy.type.gen))

##Separating into 3 different data sets to analyze different impacts
Final.gen.census <-
  full_join(processed.gen.census, total.gen.census) 

Renewable.county <- 
  Final.gen.census %>%
  subset(Renewable.v.FossilFuel=='Renewable Fuel') 

Fossil.county <- 
  Final.gen.census %>%
  subset(Renewable.v.FossilFuel=='Fossil Fuel') 
```


```{r echo=FALSE}
## Correlation Plot
Corr.data <- 
  Renewable.county %>%
  select(MedianHouseholdIncome:total.gen) %>%
  na.omit() %>%
  subset(select=-c(County))

corrplot(Corr.data, method = "ellipse")
corrplot.mixed(Corr.data, upper = "ellipse")

## All variable Regression
total.regression.variables <- 
  lm(data = Final.gen.census, log(total.gen) ~ log(MedianHouseholdIncome) + 
       Per.CollegeDeg + Per.PovertyMobility + 
       Per.MinorityMobility)
fossil.reg.variables <- 
  lm(data = Fossil.county, log(energy.type.gen) ~ log(MedianHouseholdIncome) + 
       Per.CollegeDeg + Per.PovertyMobility + 
       Per.MinorityMobility)
renewable.reg.variables <- 
  lm(data = Renewable.county, log(energy.type.gen) ~ log(MedianHouseholdIncome) + 
       Per.CollegeDeg + Per.PovertyMobility + 
       Per.MinorityMobility)

## AIC
step(regression.variables)
step(fossil.reg.variables)
step(renewable.reg.variables)

## Regression Analysis
renewable.reg <- 
  lm(data = Renewable.county, log(energy.type.gen) ~ log(MedianHouseholdIncome) + 
        Per.MinorityMobility)
summary(renewable.reg)


total.energy.reg <- 
  lm(data = Final.gen.census, log(total.gen) ~ log(MedianHouseholdIncome) + 
       Per.CollegeDeg + Per.PovertyMobility + 
       Per.MinorityMobility)
summary(total.energy.reg)

fossil.reg <- 
  lm(data = Fossil.county, log(energy.type.gen) ~ log(MedianHouseholdIncome) + 
       Per.CollegeDeg + Per.PovertyMobility + 
       Per.MinorityMobility)
summary(fossil.reg)




plot(renewable.reg)


```


For this analysis we used the generation site's nameplate capacity as the 
independent variable demonstrating the size, and assumed impact, to a community. 
We aggregated these values into three broad categories: fossil fuel, renewable, 
and total energy generation in each county. We then combined this data with 
demographic data from the US Census Bureau for the year 2020. These dependent 
variables include the county's median income, the percent of the county's total 
population that have received a college or associates degree, and the mobility
of the population based on race and poverty status. 

Given that both total energy generation within a county and their median income
are positively skewed, we used log transformation on those variables. The 
remaining dependent variables were not log transformed as we already converted 
them to percentages of each county's total population. 

To help narrow the focus of our analysis we utilized Akaike's Information 
Criterion (AIC) to select our variables for each regression. This exploratory 
analysis demonstrated that, given our data limitations, focusing on renewable
data and using our minority mobility and median income variables will offer the 
best analysis (AIC=21.31).

The regression analysis (Adj. R-squared = 0.1638) demonstrates with a 1% 
increase in a county's renewable energy infrastructure, the county's median 
income decreases by 1.51% (p-value = 0.0329) and the ratio of population that 
is a minority increases by 1.97% (p-value = 0.0117). This is in line with what 
we hypothesized, because although renewable energy is often thought of as a 
luxury resource, the infrastructure still undergoes NIMBY criticism by many 
wealthier individuals. 

## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
